Classification Model
===========================================

##Classification method for predictive modelling

###Basic settings


```{r}
echo = TRUE
setwd("c:/Users/jasmine.qi/Documents/Classification")
library(Rcmdr)
library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(party)
library(gmodels)
library(pROC)
library(rattle)
library(randomForest)
library(ROCR)
```
### Descriptive
```{r}
## Load files which in the Classification folder
total= read.csv("derive.csv")
dim(total)

#Classify datasets to Good Loan and Bad Loan

#Summary of Good loans and Bad loans
local({
  .Table <- with(total, table(Classifier))
  cat("\ncounts:\n")
  print(.Table)
  cat("\npercentages:\n")
  print(round(100*.Table/sum(.Table), 2))
})

# Histogram of Loanscores

x=total$Loanscore
h=hist(x, col = "red", breaks = 20, main = "Breaks=20 of mean", xlab=" loan performance scores" )
xfit<-seq(min(x),max(x),length=40) 
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x)) 
yfit <- yfit*diff(h$mids[1:2])*length(x) 
lines(xfit, yfit, col="blue", lwd=2)


# randomsize 70% of row of data as training set and test set
#set.seed(1)

indexes = sample(1:nrow(total), size= 0.7*nrow(total))
train = total[indexes, ]
test = total[-indexes, ]

nrow(train)
nrow(test)
```

### summary of predictors that model used

```{r}



formula=Classifier ~ Absorption.Rate...prior.4.6.months. + 
  Absorption.Rate...prior.7.12.months. + 
  Absorption.Rate...prior.current.3.months. + Appraisal.Monthly.Rent + 
  Appraisal.price + Assignment.Type + Bankrupcy + Baths + Bedrooms + 
  Borrower.Age.Group + Borrower.Rent + Built.up + Charge.offs + 
  Closed.Past.Auto.Late + Closed.Past.Education.Late + 
  Closed.Past.Installment.Late + Closed.Past.Mortgage.Late + 
  Closed.Revolving.Late + Collections + Comparable.properties.lower.price + 
  Comparable.properties.upper.price + Comparable.sale.lower.price.range + 
  Comparable.sale.upper.price.range + Condition + Credit.Score + Debt.High.Credit.... + Demand.Supply + 
  distancemiles + DOM + Expected.cost.to.Rehab63 + Expected.flip.price + 
  Foreclosure + Foreclosure.sales + 
  Full.or.Part.Time.investor + Growth + Indicated.Value.by.Cost.Approach + 
  Indicated.Value.by.Income.Approach + Interest.Rate + Lending.Value + 
  Loan.Purpose + Loan.Term + Loan.Type + Location + LTV + Male.Female + 
  Marketing.Time + Median.Comparable.List.Price.prior.4.6.months. + 
  Median.Comparable.List.Price.prior.7.12.months. + 
  Median.Comparable.List.Price.prior.current.3.months. + 
  median.Comparable.listings.Days.on.Market.prior.4.6.months. + 
  median.Comparable.listings.Days.on.Market.prior.7.12.months. + 
  median.Comparable.listings.Days.on.Market.prior.current.3.months. + 
  Median.Comparable.Sale.Price..prior.4.6.months. + 
  Median.Comparable.Sale.Price..prior.7.12.months. + 
  Median.Comparable.Sale.Price..prior.current.3.months. + 
  Median.Comparable.sales.Days.on.market.prior.4.6.months. + 
  Median.Comparable.sales.Days.on.market.prior.7.12.months. + 
  Median.Comparable.sales.Days.on.market.prior.current.3.months. + 
  Median.Sale.Price.a.s..of.List.Price.prior.4.6.months. + 
  Median.Sale.Price.a.s..of.List.Price.prior.7.12.months. + 
  Median.Sale.Price.a.s..of.List.Price.prior.current.3.months. + 
  Months.of.Housing.Supply.prior.4.6.months. + 
  Months.of.Housing.Supply.prior.7.12.months. + 
  Months.of.Housing.Supply.prior.current.3.months. + Occupant + 
  Open.Current.Auto.Late + Open.Current.Education.Late + 
  Open.Current.Installment.Late + Open.Current.Mortgage.Late + 
  Open.Current.Revolving.Late  + Origination.UPB + 
  OrigTransdays + Price.of.Prior.Sale.Transfer+ 
  Property.Values + Rent.Own + Repeat + Rooms + Sale.Price + Size + 
  Square.feet.of.gross.living.area  + 
  Subject..PropertyState + Total...of.Comarable.Sales...prior.4.6.months. + 
  Total...of.Comarable.Sales...prior.7.12.months. + 
  Total...of.Comarable.Sales...prior.current.3.months. + 
  Total...of.Comparable.Active.listings.prior.4.6.months. + 
  Total...of.Comparable.Active.listings.prior.7.12.months. + 
  Total...of.Comparable.Active.listings.prior.current.3.months. + Type + 
  Units + VFS.Internal.Value + WLV + X..of.comarable.properties + 
  X..of.comparable.sales + X..of.home.purchased.in.last.3.years + 
  X..of.Properties.currently.owned + X..of.years.investing +Yearbuilt
```

### A simple tree model-CART 
```{r}
## A simple tree model

ctrl_rpart=trainControl(method ="cv", number=10, summaryFunction = twoClassSummary, classProbs=TRUE)
train_tree= train(formula, data=train, method="rpart", tuneLength=10, metric="ROC",trControl=ctrl_rpart)

##plot tree of trainig model
fancyRpartPlot(train_tree$finalModel, cex=0.7, main = "Tree of CART Model", sub="")
plot(varImp(train_tree), top=15, xlab="Variable Imortance for CART Models")

##Predict test sets
test_pred = predict(train_tree, test)
cart=confusionMatrix(test_pred, test$Classifier)

cart
#predict class probabilities
test_predprob = predict(train_tree, test, type="prob")
head(test_predprob)

##plot ROC curve
test_predroc = roc(response=test$Classifier, predictor = test_predprob$Bad, levels = rev(levels(test$Classifier)))




```
### C5 model
```{r}

##c5
grid = expand.grid(.model="tree",
                   .trials=c(1:100),
                   .winnow=F)
train_treec5 = train(formula,data=train, method="C5.0", metric="ROC", tuneGrid=grid, trControl=ctrl_rpart)



## plot c5 model Tree
plot(train_treec5)
## predict test datasets and Accurarcy

test_predc5= predict(train_treec5, test)
c5=confusionMatrix(test_predc5, test$Classifier)

c5
#Test set ROC Curve

test_predc5prob = predict(train_treec5, test, type = "prob")
head(test_predc5prob)

test_predc5roc = roc(response=test$Classifier, predictor = test_predc5prob$Bad, levels = rev(levels(test$Classifier)))

```


##Random Forest method

```{r}

#set.seed(1)

#Regardless unbalanced class
train_rf=train(formula, data=train, method="rf", trControl=ctrl_rpart, prox=T, allowParallel=T, metric="ROC")

#Consider unbalanced class - Down sampling Good loans
train_rf2=randomForest(formula, data=train, ntree=train_rf$finalModel$ntree, mtry=train_rf$finalModel$mtry, proximity=T, strata=train$Classifier,sampsize=rep(min(table(train$Classifier)), nlevels(train$Classifier)))


#Unbalanced
train_rf$finalModel

#Balanced
train_rf2

## prediction test data

test_predrf=predict(train_rf, newdata=test) 
test_predrf2=predict(train_rf2, test) 

#Unbalanced
cftable_rf=CrossTable(test$Classifier, test_predrf, digits=1, prop.r = F, prop.t = F, prop.chisq = F, chisq = F, data=test)
table_rf =table(test$Classifier, test_predrf)

## Accuracy of test unbalanced class
rf=confusionMatrix(test_predrf, test$Classifier)
#Balanced

rf2=confusionMatrix(test_predrf2, test$Classifier)

## plot ROC curve

## Test sets ROC curve
test_predrfprob =  predict(train_rf, test, type = "prob")
head(test_predrfprob)
test_predrfroc = roc(response=test$Classifier, predictor = test_predrfprob$Bad, levels = rev(levels(test$Classifier)))

test_predrf2prob =  predict(train_rf2, test, type = "prob")
head(test_predrf2prob)
test_predrf2roc = roc(response=test$Classifier, predictor = test_predrfprob$Bad, levels = rev(levels(test$Classifier)))





##plot importance variables
#importance(train_rf)
par(mfrow=c(1,2))
plot(varImp(train_rf), top=15, main="Unbalanced RandomFOrest")
varImpPlot(train_rf2, n=15, main="Balanced RandomFOrest")
```

##GBM Gradient Boosting model
```{r}


train_gbm = train(formula, data=train, method="gbm", trControl=ctrl_rpart, metric="ROC", maximize=F, verbose=F)
plot(train_gbm)


##importance virables for GBM model
gbmImp = varImp(train_gbm, scale=F)
plot(gbmImp, top=15, main="GBM model")

#predict testdatasets
test_gbm=predict(train_gbm, test)
cftable_gbm=CrossTable(test$Classifier, test_gbm, digits=1, prop.r = F, prop.t = F, prop.chisq = F, chisq = F, data=test)
table_gbm =table(test$Classifier, test_gbm)

#Accuracy of Gradient Bossting model
gbm=confusionMatrix(test_gbm, test$Classifier)

#ROC figure
test_predgbmprob =  predict(train_gbm, test, type = "prob")
head(test_predgbmprob)

test_predgbmroc = roc(response=test$Classifier, predictor = test_predgbmprob$Bad, levels = rev(levels(test$Classifier)))





```



## Summary of all the modeling
###ROC
```{r}
resamps= resamples(list(CART=train_tree, C5=train_treec5, rfUnbalance=train_rf, GBM=train_gbm))

resamps
# all the performence results
summary(resamps)

# Figures of comparison of six models

bwplot(resamps,layout=c(3,1))

dotplot(resamps, metric="ROC")

```
C5 model covers more in ROC figures, GBM has less sensitivity.

###Accuracy table for Models


```{r}

accuracy1=as.data.frame(cart$overall[1])
colnames(accuracy1)="CART"
accuracy2=as.data.frame(c5$overall[1])
colnames(accuracy2)="C5"
accuracy3=as.data.frame(rf$overall[1])
colnames(accuracy3)="Unbalanced RF"
accuracy4=as.data.frame(rf2$overall[1])
colnames(accuracy4)="balanced RF"
accuracy5=as.data.frame(gbm$overall[1])
colnames(accuracy5)="GBM"
accuracy=cbind(accuracy1, accuracy2, accuracy3, accuracy4, accuracy5)


accuracy
```

Unbalanced Random Forest has the highest accuracy, but the Blanced Random Forest has the lowest accuracy.

### Sensitivity of modeling fitting Comparison

```{r}
plot(test_predroc, type="S", lty= 1, col="red", print.thres = 0.5, main="Classification ROC Comparison")
plot(test_predc5roc,lty=2, add = T, col="green")
plot(test_predrfroc, add= T, col="orange", lty=3)
plot(test_predrf2roc, add = T, col="blue", lty=4)
plot(test_predgbmroc, col="pink", add=T, lty=5)
legend(0.6, 0.6, c("CART Tree", "C5 Tree", "Unbalanced Random Forest", "Balanced Random Forest", "GBM"), col=c("red", "green", "orange", "blue", "pink"), lwd = 5)
```